Contracts for Contracts

### 2023-05-08 paper/introduction.tex
intro

Smart contracts are programs to automate the execution of agreement so that all
parties involved in the agreement can be certain about the outcome.
Smart contracts are ``smart'' since by using blockchains they provide a
mechanism to \emph{ensure the economic outcome of transactions} without a
central authority.
A blockchain is a \note{XXX} that implements a consensus protocol of
Byzantine-fault-tolerant distributed ledgers.

It is however ironic that many popular smart contract languages, which hold the
great promise to ensure transactional outcomes via consensus and distributed
ledgers, do not provide enough means to \emph{ensure the computational outcomes},
leading to poor quality of programs.
Without ensuring the quality of programs, it is impossible to ensure
the economic outcome of transactions.

Indeed, we have seen a great number of notorious attacks to smart contract
systems in recent years.
For example, \todo{explain attack using callback}.
Attackers are able to utilize
the callback mechanism to \note{do crazy things and steal money}
The root cause of this attack is \todo{explain}.

Solidity does provide first-order assertion primitives (e.g. @require@) but
they cannot effectively check function values, which are a common pattern used
in smart contract programs.
\zz{
On the other hand, assertion primitives are crucial for ensuring the accuracy and
integrity of a smart contract's business model.
}
\zz{
I'm not sure if this statement applies to the PL community, but from my
understanding, assertions are typically used during testing rather than runtime.
However, in smart contracts, these assertions function as guard conditions and can cause reverts if violated, making them extremely important.
}
Programmers have to manually insert such checks around relevant call-sites, \todo{which is bad}.
%Solidity also provides the @modifier@ mechanism that intercepts
%before and after a function call, thus can be used to check
%pre- and post-conditions.
%\todo{can we compose modifier?}
%\todo{can we attach modifier to callbacks?}
%\todo{can we use modifier to check function's argument?}

The lack of sufficient linguistic means to specify and enforce subtle behaviors
discourages programmers to use concise higher-level abstractions when writing
smart contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level abstractions, which conflates the main business logic and
behavioral checks, leading to poor readability and maintainability.
For instance, to defend the famous DAO attack \note{cite} caused by allowing
``reentrancy'' via external functions, programmers have to be extra cautious
about the execution order or maintain additional states to check if a call is
allowed. Either way is not ideal.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

Apart from investigating effort to audit, certify, or verify smart contract programs,
language designers are responsible to empower programmers with better
linguistic solutions to secure smart contracts in the first place.
In this paper, we show that advanced \emph{behavioral software contracts} provide a
lightweight solution for securing \emph{smart contracts}.

>>>>

Smart contracts are programs to automate the execution of agreement so that all
parties involved in the agreement can be certain about the outcome.
Smart contracts are ``smart'' since by using blockchains they provide a
mechanism to \emph{ensure the economic outcome of transactions} without a
central authority.
A blockchain implements a consensus protocol of Byzantine-fault-tolerant
distributed ledgers.

It is however ironic that many popular smart contract languages, which hold the
great promise to ensure transactional outcomes via consensus and distributed
ledgers, do not provide enough means to \emph{ensure the computational outcomes}.
This unfortunate reality leads to poor quality of programs.
Without ensuring the quality of programs, it is impossible to ensure the
economic outcome of transactions.

Indeed, in recent years, we have seen a number of failures of smart contract
systems due to attacks utilizing vulnerable contract programs.
For example, in the famous DAO hack\note{cite}, the attacker is able to utilize the
callback mechanism to create recursive function calls and temporarily stole
money, causing a total lose of \$50 million.
The root cause of this attack is that the contract does not deploy defensive
checks before performing critical actions.

%\zz{ On the other hand, assertion primitives are crucial for ensuring the
%accuracy and integrity of a smart contract's business model. }
%\zz{ I'm not sure if this statement applies to the PL community, but from my
%understanding, assertions are typically used during testing rather than
%runtime.  However, in smart contracts, these assertions function as guard
%conditions and can cause reverts if violated, making them extremely important. }

%Solidity also provides the @modifier@ mechanism that intercepts
%before and after a function call, thus can be used to check
%pre- and post-conditions.
%\todo{can we compose modifier?}
%\todo{can we attach modifier to callbacks?}
%\todo{can we use modifier to check function's argument?}

We observe that popular smart contract languages (such as Solidity) do
not provide \emph{rich, effective, convenient} means to specify and enforce
contract behaviors.
Simple first-order assertion primitives (e.g., @require@) are indeed common,
but they cannot effectively check callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical behavior in smart contracts are \emph{temporal}.
For instance, the defense to the DAO attack \note{cite} is to
prevent ``reentrancy'' via external functions, which requires that the function
of interest cannot be invoked \emph{before} the current call of this function is finished.

The lack of sufficient linguistic means to specify and enforce subtle but
critical behaviors discourages programmers to use concise higher-level
abstractions when writing smart contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level or no abstractions, which conflates the main business logic with
behavioral checks, further leading to poor readability and maintainability.
For example, a common way to implement ``reentrancy guards'' is to manually
make an additional boolean state, marking if the function is finished.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

Apart from the investigating effort to certify and verify smart contract
programs, language designers are responsible to empower programmers with better
linguistic solutions to secure smart contracts in the first place.
In this paper, we develop practical \emph{higher-order and temporal behavioral
software contracts} that provide a lightweight solution for securing
\emph{smart contracts}.
